---
title: "Getting started with sparsefusion"
author: "Daniel Agyapong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with sparsefusion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `sparsefusion` package fits grouped fused-regression models. The data
interface is a feature matrix `X`, response vector `y`, group labels `groups`,
and a fusion graph matrix `G` whose nonzero entries define which groups are
encouraged to share regression coefficients.

The recommended entry point is `sparse_fusion()`. Use `fusion = "l1"` for the
smoothed L1 fused solver and `fusion = "l2"` for the augmented-design L2 solver.
The default solver is `active_edge`, which works directly with the nonzero
edges of `G`.

## Simulated grouped data

```{r}
library(sparsefusion)

set.seed(1)

k <- 4L
p <- 8L
n_group <- 10L
groups <- rep(seq_len(k), each = n_group)

X <- matrix(rnorm(length(groups) * p), nrow = length(groups), ncol = p)

beta <- matrix(0, nrow = p, ncol = k)
beta[1:2, 1:2] <- 1
beta[1:2, 3:4] <- -1

y <- rowSums(X * t(beta[, groups, drop = FALSE])) +
  rnorm(length(groups), sd = 0.1)
```

The matrix `beta` has one coefficient vector for each group. Here the first two
groups share one coefficient pattern and the last two groups share another.

## Sparse fusion graph

The fusion graph is represented as a symmetric matrix over groups. This example
uses a chain graph.

```{r}
G <- matrix(0, k, k)
for (i in seq_len(k - 1L)) {
  G[i, i + 1L] <- 1
  G[i + 1L, i] <- 1
}

G
```

## Fit an L1 active-edge model

```{r}
fit <- sparse_fusion(
  X, y, groups, G,
  lambda = 1e-3,
  gamma = 1e-2,
  fusion = "l1",
  solver = "active_edge",
  mu = 1e-4,
  tol = 1e-3,
  num.it = 800,
  intercept = FALSE,
  scaling = FALSE
)

dim(fit)
```

The result is a coefficient matrix with one column per group.

## Predict on new observations

For observations whose groups were present during fitting, predictions are
obtained by selecting the corresponding coefficient column.

```{r}
X_new <- X[1:6, , drop = FALSE]
groups_new <- groups[1:6]

y_hat <- numeric(nrow(X_new))
for (g in unique(groups_new)) {
  idx <- groups_new == g
  y_hat[idx] <- X_new[idx, , drop = FALSE] %*% fit[, g]
}

y_hat
```

## Solver modes

The `solver` argument controls the graph representation used by the optimizer.

```{r, eval=FALSE}
# Exact sparse graph representation, recommended default.
sparse_fusion(X, y, groups, G, lambda, gamma,
  fusion = "l1", solver = "active_edge"
)

# Full all-pair reference construction.
sparse_fusion(X, y, groups, G, lambda, gamma,
  fusion = "l1", solver = "full_pairwise"
)

# Approximate L1 chain solver.
sparse_fusion(X, y, groups, G, lambda, gamma,
  fusion = "l1", solver = "chain_approx"
)
```

The same interface is used for L2 fusion. The `chain_approx` solver is only
available for L1 fusion.

```{r}
fit_l2 <- sparse_fusion(
  X, y, groups, G,
  lambda = 1e-3,
  gamma = 1e-2,
  fusion = "l2",
  solver = "active_edge",
  intercept = FALSE,
  scaling = FALSE
)

dim(fit_l2)
```
